Drawing Math

### 2022-05-13 paper/Paper.tex
clean

\section{Introduction}
A glum Pittsburgh day inspired the authors to take a random walk down YouTube's recommended
when they encountered the Numberphile video on ``Plotting Pi" \cite{NumberphileVid}.
In the video, Henderson and Brady introduce the idea of taking a Python Turtle and
deriving a series of updates to its position based off of various decimal sequences, 
some rational, some irrational. More specifically, they place an object in 2D, cartesian
space starting at $(0, 0)$ and ``facing" to the right. They then take a generating number, like $\pi$ or $35/99$, 
which gives a decimal sequence ($3.14...$ and $0.\bar{35}$ respectively). 
Then, at time step $i$, $i \geq 1$, they rotate the object by the $i$th digit divided by the base of
the decimal sequence. Then, the Turtle moves a constant distance in the direction which it faces.
For example, if the second digit is 4, the turtle is rotated by $\frac{4}{10}$ths of a circle
counterclockwise at time step 2 and then moves 10 units in its new direction.
\begin{figure}[ht]
  \centering
  \includegraphics[width=8cm]{fig/turtle2dtimestep.png}
  \caption{The turtle moving according to decimal sequence $0.223$ in base $10$.}
\end{figure}

Whenever the Turtle moves, the program draws a red line along the Turtle's path. Thus,
different lines are drawn.
The authors noticed that using rational numbers to generate the sequence often drew geometrically
aesthetic, closed shapes.

\begin{figure}[ht]
  \centering
  \includegraphics[width=8cm]{fig/Turtle1D13Div113.png}
  \caption{The closed shape generated from $\frac{13}{113} = 0.0977443609...$ in base $10$.}
\end{figure}
\begin{figure}[ht]
  \centering
  \includegraphics[width=8cm]{fig/Turtle1DNotClosed.png}
  \caption{A non-closing shape generated from $\frac{2134}{9999} = 0.\overline{2134}$ in base $10$.}
\end{figure}

Naturally, the authors were curious whether a Turtle in ``3D" would also draw 
aesthetic shapes. In other words, what if the Turtle's orientation was described
by 2 angles, pitch and yaw. The pitch and yaw would then by independently updated
by 2 decimal sequences generated from 2 rational numbers.
% TODO: figure
After writing the program, we noticed that the Turtle would often fail
to close. In other words, the Turtle would go off in one direction forever.
% TODO: picture of failing to close.
But, the Turtle \textit{sometimes} closed.

The authors then proceeded to ask why the Turtle would sometimes close and sometimes
go off into the ether. After finding a closed form solution in 3 dimensional space,
we proceeded to ask the same question in $D$ dimensional space: given $D - 1$ 
rational decimal sequences which determine the Turtle's orientation in space,
does the Turtle move in a closed shape (i.e.\ is the Turtle position always contained
within some $D$ dimensional sphere off constant radius)?

@@

\section{Definitions and questions}
\subsection{Definitions}
Say you (yes you!) had a turtle living in $D$ dimensional Euclidean
space and in discrete time. At time step $i$, where $i \in \Z$ and $i > 0$,
the turtle has position $p_i \in \R^D$.
Then, lets define $\Delta p_{i+1} = p_{i+1} - p_i$; in other words, $\Delta p_{i+1}$ is the change in position from time $i$ to $i + 1$.

@@

\section{Does it close?}
In understanding whether a set of given rationals, bases, and updated functions 
draw a closed shape in $D$ dimensional space, we first need to find the period of 
the update delta, $\Delta p_i$. We then know that the total update over a period will be repeated 
indefinitely. Consequently, we then seek to find the total change in position over a period. 
If the total change is $0$, the shape will close as the Turtle will end up at its starting point
after every period length. If the total update is nonzero, the Turtle will not draw a closed
shape.

@@

Then, let
$$
  \mathcal{B} = \left\{\beta: f_\beta(\sigma) = 0\right\}
$$
and $\overline{\mathcal{B}} = \mathcal{M} \setminus \mathcal{B}$. Then, if for all
$\beta \in \mathcal{B}$,
$$
\sum_{q = 0}^{\rationalPeriodTotal - 1}
  \exp\left(
      I 
      \commonToAngle 
      \sum_{\ell = 0}^{q}
      \sum_{j \in A_d}
          \left(-1\right) ^ {\beta_{j}}
          \numbToCommon
          \mathrm{digit}(n, b, d, \ell)
  \right) = 0
$$
the Turtle will draw a closed shape. While this may seem arbitrary, there is a unique
and interesting geometric interpretation. 
Let a rational decimal sequence $a_1, a_2, ..., a_{\rationalPeriodTotal}$ be equal to
$
\forall \ell \in [\rationalPeriodTotal]
$,
$$
a_\ell = \sum_{j \in A_d}
\left(-1\right) ^ {\beta_{j}}
\numbToCommon
\mathrm{digit}(n, b, d, \ell).
$$
In other words, we are creating a ``common" rational sequence by summing and subtracting
our rational parameters in the least common multiple base $\commonBase$. Then,
if $$
\sum_{q = 0}^{\rationalPeriodTotal - 1}
\exp\left(
    I 
    \commonToAngle 
    \sum_{\ell = 0}^{q}
      a_\ell
  \right) = 0
$$
the Turtle closes.
If $r_i = \commonToAngle \sum_{\ell = 0}^{q} a_\ell$ for 
$i \in \set{0, ..., \rationalPeriodTotal - 1}$, then $\exp(I \cdot r_i)$ can be thought of as some point at angle $r_i$ around the 
complex unit circle. Also, note that $r_{i + 1} = r_i + \commonToAngle a_{i+1}$. In otherwords,
$r_i$ denotes some position around the complex unit circle where the subsequent updates
to position are given the $a_i$s, which are a sum and subtraction of rational parameters. Then, if the center of
mass of these points around the unit circle is $0$, the Turtle draws a closed shape!
The authors are unsure as to the underlying intuition for why this is true.

>>>>

\section{Introduction}
A glum Pittsburgh day inspired the authors to take a random walk down YouTube's recommended
when they encountered the Numberphile video on ``Plotting Pi" \cite{NumberphileVid}.
In the video, Henderson and Brady introduce the idea of taking a Python Turtle and
deriving a series of updates to its position based off of various decimal sequences, 
some rational, some irrational. More specifically, they place an object in 2D, cartesian
space starting at $(0, 0)$ and ``facing" to the right. They then take a generating number, like $\pi$ or $35/99$, 
which gives a decimal sequence ($3.14...$ and $0.\bar{35}$ respectively). 
Then, at time step $i$, $i \geq 1$, they rotate the object by the $i$th digit divided by the base of
the decimal sequence. Then, the turtle moves a constant distance in the direction which it faces.
For example, if the second digit is 4, the turtle is rotated by $\frac{4}{10}$ths of a circle
counterclockwise at time step 2 and then moves 10 units in its new direction.
\begin{figure}[h]
  \centering
  \includegraphics[width=8cm]{fig/turtle2dtimestep.png}
  \caption{The turtle moving according to decimal sequence $0.223$ in base $10$.}
\end{figure}

Whenever the turtle moves, the program draws a red line along the turtle's path. Thus,
different lines are drawn.
The authors noticed that using rational numbers to generate the sequence often drew geometrically
aesthetic, closed shapes.

\begin{figure}[h]
  \centering
  \includegraphics[width=8cm]{fig/Turtle1D13Div113.png}
  \caption{The closed shape generated from $\frac{13}{113} = 0.0977443609...$ in base $10$.}
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics[width=8cm]{fig/Turtle1DNotClosed.png}
  \caption{A non-closing shape generated from $\frac{2134}{9999} = 0.\overline{2134}$ in base $10$.}
\end{figure}

Naturally, the authors were curious whether a Turtle in ``3D" would also draw 
aesthetic shapes. In other words, what if the turtle's orientation was described
by 2 angles, pitch and yaw. The pitch and yaw would then by independently updated
by 2 decimal sequences generated from 2 rational numbers.
% TODO: figure
\begin{figure}[h]
  \centering
  \includegraphics[width=8cm]{fig/turtle.png}
  \caption{The above is a placeholder, I would like to get Shiva to help me design a 3D graphic here}
\end{figure}
After writing the program, we noticed that the turtle would often fail
to close. In other words, the turtle would go off in one direction forever.
% TODO: picture of failing to close.
But, the turtle \textit{sometimes} closed.

The authors then proceeded to ask why the turtle would sometimes close and sometimes
go off into the ether. After finding a closed form solution in 3 dimensional space,
we proceeded to ask the same question in $D$ dimensional space: given $D - 1$ 
rational decimal sequences which determine the turtle's orientation in space,
does the turtle move in a closed shape (i.e.\ is the turtle position always contained
within some $D$ dimensional sphere off constant radius)?

@@

\pagebreak
\section{Definitions and questions}
\subsection{Definitions}
Say we had a turtle, Jimbo, living in $D$ dimensional Euclidean
space and in discrete time. At time step $i$, where $i \in \Z$ and $i > 0$,
the turtle has position $p_i \in \R^D$.
Then, lets define $\Delta p_{i+1} = p_{i+1} - p_i$; in other words, $\Delta p_{i+1}$ is the change in position from time $i$ to $i + 1$.

@@

\section{Does it close?}
In understanding whether a set of given rationals, bases, and updated functions 
draw a closed shape in $D$ dimensional space, we first need to find the period of 
the update delta, $\Delta p_i$. We then know that the total update over a period will be repeated 
indefinitely. Consequently, we then seek to find the total change in position over a period. 
If the total change is $0$, the shape will close as the turtle will end up at its starting point
after every period length. If the total update is nonzero, the turtle will not draw a closed
shape.

@@

Then, let
$$
  \mathcal{B} = \left\{\beta: f_\beta(\sigma) = 0\right\}
$$
and $\overline{\mathcal{B}} = \mathcal{M} \setminus \mathcal{B}$. Then, if for all
$\beta \in \mathcal{B}$,
$$
\sum_{q = 0}^{\rationalPeriodTotal - 1}
  \exp\left(
      I 
      \commonToAngle 
      \sum_{\ell = 0}^{q}
      \sum_{j \in A_d}
          \left(-1\right) ^ {\beta_{j}}
          \numbToCommon
          \mathrm{digit}(n, b, d, \ell)
  \right) = 0
$$
the turtle will draw a closed shape. While this may seem arbitrary, there is a unique
and interesting geometric interpretation. 
Let a rational decimal sequence $a_1, a_2, ..., a_{\rationalPeriodTotal}$ be equal to
$
\forall \ell \in [\rationalPeriodTotal]
$,
$$
a_\ell = \sum_{j \in A_d}
\left(-1\right) ^ {\beta_{j}}
\numbToCommon
\mathrm{digit}(n, b, d, \ell).
$$
In other words, we are creating a ``common" rational sequence by summing and subtracting
our rational parameters in the least common multiple base $\commonBase$. Then,
if $$
\sum_{q = 0}^{\rationalPeriodTotal - 1}
\exp\left(
    I 
    \commonToAngle 
    \sum_{\ell = 0}^{q}
      a_\ell
  \right) = 0
$$
the turtle closes.
If $r_i = \commonToAngle \sum_{\ell = 0}^{q} a_\ell$ for 
$i \in \set{0, ..., \rationalPeriodTotal - 1}$, then $\exp(I \cdot r_i)$ can be thought of as some point at angle $r_i$ around the 
complex unit circle. Also, note that $r_{i + 1} = r_i + \commonToAngle a_{i+1}$. In otherwords,
$r_i$ denotes some position around the complex unit circle where the subsequent updates
to position are given the $a_i$s, which are a sum and subtraction of rational parameters. Then, if the center of
mass of these points around the unit circle is $0$, the turtle draws a closed shape!
The authors are unsure as to the underlying intuition for why this is true.
